The \dealii{} Library, Version 9.2

### 2020-06-01 9.2/paper.tex
Minor edits.

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  for a while already
  (e.g., ones in quantum mechanics -- like the equation used in the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems). However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
\item The new \texttt{DiscreteTime} class provides a more
  consistent, more readable, and less error-prone approach to control
  time-stepping algorithms within time-dependent simulations.
  The interface of this class is designed to be minimal
  to enforce a number of important programming invariants, reducing
  the possibility of mistakes in the user code.
  An example is that \texttt{DiscreteTime} ensures that the final time step ends
  precisely on a predefined end time, automatically
  lengthening or shortening to final time step.
\item A key component of \dealii{} are the \texttt{FEValues} and
  \texttt{FEFaceValues} classes that evaluate finite element functions
  at quadrature points located on cells and faces of a cell,
  respectively~\cite{BangerthHartmannKanschat2007}. This release now
  contains a class \texttt{FEInterfaceValues} that considers the
  restriction of the shape functions from \textit{both} sides of a face and
  allows evaluating jumps and averages of shape functions along this
  face. These are common components of the bilinear forms of
  discontinuous Galerkin schemes (as well as schemes for fourth-order
  equations, see the discussion of \texttt{step-47} below) and greatly
  simplify the implementation of these methods.
\item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
  class, which transfers local quadrature point data onto the
  children of newly refined cells, did not allow different regions of the
  triangulation store quadrature data of different lengths. This release lifts
  this restriction, which enables efficient quadrature data storage and data
  transfer within a triangulation containing multiple material models, each with
  their own number of local state variables and history variables.
  As an example, in a solid mechanics simulation, we can assign a hyper-elastic
  material model can be assigned to a region of the mesh with no associated
  history variables while in another part of the triangulation we incorporate an
  elasto-plastic constitutive model, requiring the storage of local plasticity
  data at quadrature points.
  During each refinement of the mesh, we can use
  \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
  plasticity data from parent cells to their children, ignoring the cells
  with the hyper-elastic constitutive model.
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

@@

\item \texttt{step-69} presents a first-order scheme solving compressible
Euler's equations of gas dynamics with a graph-viscosity stabilization
technique. Beside the usual conservation properties of mass, momentum and
total energy, the method also guarantees that the constructed solution
obeys pointwise stability constraints (positivity of density, internal
energy and a local minimum principle on the specific entropy). As such
\text{step-69} is strictly speaking more a collocation-type discretization
than a variational formulation, even though it is implemented with finite
elements.

The time-update at each node requires the evaluation of a right-hand side
that depends (nonlinearly) on information from the previous time-step that
spans more than one cell. Therefore, assembly loops operate directly on the
sparsity graph in order to retrieve information from the entire stencil
associated to each node. From a programming perspective, \texttt{step-69}
features a number of techniques that are of interest for a wider audience:
It discusses a hybrid thread and MPI parallelized scheme with efficient
(MPI node) local numbering of degrees of freedom. It showcases how to
perform asynchronous write-out of results using a background thread with
\texttt{std::async}, and discusses a simple but effective checkpointing and
restart technique.

>>>>

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  for a while already
  (e.g., ones in quantum mechanics -- like the equation used in the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems). However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
\item The new \texttt{DiscreteTime} class provides a more
  consistent, more readable, and less error-prone approach to control
  time-stepping algorithms within time-dependent simulations.
  The interface of this class is designed to be minimal
  to enforce a number of important programming invariants, reducing
  the possibility of mistakes in the user code.
  An example is that \texttt{DiscreteTime} ensures that the final time step ends
  precisely on a predefined end time, automatically
  lengthening or shortening to final time step.
\item A key component of \dealii{} are the \texttt{FEValues} and
  \texttt{FEFaceValues} classes that evaluate finite element functions
  at quadrature points located on cells and faces of a cell,
  respectively~\cite{BangerthHartmannKanschat2007}. This release now
  contains a class \texttt{FEInterfaceValues} that considers the
  restriction of the shape functions from \textit{both} sides of a face and
  allows evaluating jumps and averages of shape functions along this
  face. These are common components of the bilinear forms of
  discontinuous Galerkin schemes (as well as schemes for fourth-order
  equations, see the discussion of \texttt{step-47} below) and greatly
  simplify the implementation of these methods.
\item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
  class, which transfers local quadrature point data onto the
  children of newly refined cells, did not allow different cells
  to store quadrature data of different lengths. This release lifts
  this restriction, which enables efficient quadrature data storage and data
  transfer within a triangulation containing multiple material models, each with
  their own number of local state variables and history variables.
  As an example, in a solid mechanics simulation, we can assign a hyper-elastic
  material model to a region of the mesh with no associated
  history variables, while in another part of the triangulation we incorporate an
  elasto-plastic constitutive model, requiring the storage of local plasticity
  data at quadrature points.
  During each refinement of the mesh, we can then use
  \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
  plasticity data from parent cells to their children, ignoring the cells
  with the hyper-elastic constitutive model.
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

@@

Support for particles was originally introduced in \dealii{} version 9.0. These
particles can be used as passive tracers, or as part of more complex
models such as those based on Particle-In-Cells (PIC) approaches \cite{GLHPB18}.

With the current release, support for particles has been further expanded: New parallel
insertion mechanisms and a basic interface to post-process particles have
been added, effectively making their usage more flexible and enabling a larger
range of use cases (such as the immersed boundaries in step-70, see below).

Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles()} member function,
particles can now be inserted in parallel from a vector of points even if these points do not lie
on the subdomain from which the insertion is called. This operation requires extensive 
communication between the processes to locate   the MPI process that
owns the cell in which the particle
is located. However, it is made significantly faster  through the usage of bounding boxes
that surround the cells of each subdomain. 
This function also takes care of transferring the properties attached to the particles to their new owner.
This new capability enables particle generators that insert particles at the location
of the support points (\texttt{Particles::Generator::dof\_support\_points()}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points()}) of a
possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
using unstructured grids generated outside of \dealii{}.

To visualize the motion of particles, the \texttt{Particles::DataOut} class was added to the library.

@@

\item \texttt{step-69} presents a first-order scheme solving the compressible
Euler equations of gas dynamics with a graph-viscosity stabilization
technique. Beside the usual conservation properties of mass, momentum, and
total energy, the method also guarantees that the constructed solution
obeys pointwise stability constraints (in particular positivity of
density and internal
energy, and a local minimum principle on the specific entropy). As such
\text{step-69} is strictly speaking more a collocation-type discretization
than a variational formulation, even though it is implemented with finite
elements.

The time-update at each node requires the evaluation of a right-hand side
that depends (nonlinearly) on information from the previous time-step that
spans more than one cell. Therefore, assembly loops operate directly on the
sparsity graph in order to retrieve information from the entire stencil
associated with each node. From a programming perspective, \texttt{step-69}
features a number of techniques that are of interest for a wider audience:
It discusses a hybrid thread and MPI parallelized scheme with efficient
MPI node-local numbering of degrees of freedom. It showcases how to
perform asynchronous write-out of results using a background thread with
\texttt{std::async}, and discusses a simple but effective checkpointing and
restart technique.
